hbm_paper_results

### 2024-02-23 paper_lbcc_results.Rmd
looking at CT sigmas, plots

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

@@

Sample plot
```{r}
#reorder studies by age
age.df <- cent_deltas_df %>%
  group_by(study) %>%
  summarize(mean_age = mean(age_years)) %>%
  arrange(mean_age)

@@

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=sex_prop_Female)) +
  geom_point() +
  geom_smooth(method="lm")
```

##Feature-wise
do deltas differ by feature?
```{r}
cent_deltas_df_long2 <- cent_deltas_df_long %>%
   mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

cent_deltas_df_long2 %>%
  ggplot() +
  geom_histogram(aes(x=delta, fill=pheno), position="identity", alpha=0.3) +
  theme(legend.position="none")

cent_deltas_df_long2 %>%
  ggplot() +
  geom_histogram(aes(x=delta*100, fill=pheno_cat, color=pheno_cat), position="identity", alpha=0.7, bins=60)

cent_deltas_df_long2 %>%
  ggplot() +
  geom_histogram(aes(x=delta*100, fill=study, color=study), position="identity", alpha=0.3, bins=60) +
  facet_wrap(~pheno_cat, scales="free") +
  theme(legend.position="none")
```

#forward selection for site_mean_abs.delta
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
full.delta.age.fit <- lm((site_mean_abs.delta*100) ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
sjPlot::plot_model(full.delta.age.fit, type = "pred", terms = "site_mean_age_yrs")
```

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")

age_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores")

# skim(delta_sum.df_piv) #some missing data! check on this!
```

```{r}
# %female
summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
full.delta.sex.fit <- lm((100*site_mean_delta) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot

```{r}
delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site))

sex_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=(site_mean_delta*100))) +
  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
  geom_smooth(method="lm", color="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores")

## Plots

#Site Effects

@@

#FIGURE 4

```{r}
# study demographics
fig5 <- ggarrange(
ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          # align = "hv", 
          common.legend = TRUE,
          legend = "bottom",
          widths=c(2, 1),
          heights=c(14, 1)),
# mean effects
ggarrange(age_effects.plt,
sex_effects.plt,
nrow=2, ncol=1,
align="hv",
labels=c("B", "C")),
common.legend=FALSE,
nrow=1, ncol=2,
labels=c("A", NA),
widths=c(1, 1.3)
)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

@@

Sample plot
```{r}
#reorder studies by age
age.df <- cent_deltas_df %>%
  group_by(study) %>%
  summarize(mean_age = mean(age_years)) %>%
  arrange(-mean_age)

@@

delta_sum.df_merged %>%
  ggplot(aes(y=site_max_abs.delta*100, x=sex_prop_Female)) +
  geom_point() +
  geom_smooth(method="lm")
```
#Site Effects

@@

##Age

#forward selection for site_mean_abs.delta
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged)) #small effect but remains significant/grows stronger
full.delta.age.fit <- lm(site_mean_abs.delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space

full.delta.age.fit2 <- lm(site_mean_delta ~ site_mean_age_yrs + site_n + sex_prop_Female, data=delta_sum.df_merged) #scaling to centile space
summary(full.delta.age.fit2)
```

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=site_mean_delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")
delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=site_mean_abs.delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")

age_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
age_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]

age_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
  theme(axis.title.x=element_b)

age_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
age_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]

age_effects.plt.null <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  annotate("text", x=60, y=-0.3, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Difference in Centile Scores")

#add prediction to df
delta_sum.df_merged$abs.delta_pred <- predict(full.delta.age.fit)

full.delta.age.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$x,
                                    predicted=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$predicted,
                                    conf.low=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.low,
                                    conf.high=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.high)
full.delta.age.fit.df <- full.delta.age.fit.df %>%
  mutate(predicted = predicted*100,
         conf.low=conf.low*100,
         conf.high=conf.high*100)

marg.age_effects.plt <- ggplot() +
  geom_count(data=delta_sum.df_merged, aes(x=site_mean_age_yrs, y=(abs.delta_pred*100), size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_line(aes(x=full.delta.age.fit.df$x, y = full.delta.age.fit.df$predicted), size = 0.7, color="#2A5783FF") +
  geom_ribbon(aes(x=full.delta.age.fit.df$x,y= NULL, ymin= full.delta.age.fit.df$conf.low, ymax= full.delta.age.fit.df$conf.high), alpha=0.1, fill="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Mean Age (Years)", y = "Predicted Mean Absolute Difference in Centile Scores")
```

##Prop Female

```{r}
# %female
summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
full.delta.sex.fit <- lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot

```{r}
delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site))
```

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")
delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="gam")

sex_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
sex_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]

sex_effects.plt <- delta_sum.df_merged %>%
  mutate(site_mean_delta=site_mean_delta*100) %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
  geom_smooth(method="lm",
              color="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores") +
  theme(axis.title.y=element_blank()) +
  annotate("text", x=0.25, y=-0.3, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
  guides(size = "none")

sex_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
sex_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]

sex_effects.plt.null <- delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
  geom_smooth(method="lm", color="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N") +
    annotate("text", x=0.25, y=4, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
  theme(axis.title=element_blank()) +
  guides(size = "none")

#add prediction to df
delta_sum.df_merged$delta_pred <- predict(full.delta.sex.fit)

full.delta.sex.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$x,
                                    predicted=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$predicted,
                                    conf.low=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.low,
                                    conf.high=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.high)
full.delta.sex.fit.df <- full.delta.sex.fit.df %>%
  mutate(predicted = predicted*100,
         conf.low=conf.low*100,
         conf.high=conf.high*100)

marg.sex_effects.plt <- ggplot() +
  geom_count(data=delta_sum.df_merged, aes(x=sex_prop_Female, y=(delta_pred*100), size=site_n, color=site_mean_age_yrs)) +
  geom_line(aes(x=full.delta.sex.fit.df$x, y = full.delta.sex.fit.df$predicted), size = 0.7, color="#24693DFF") +
  geom_ribbon(aes(x=full.delta.sex.fit.df$x,y= NULL, ymin= full.delta.sex.fit.df$conf.low, ymax= full.delta.sex.fit.df$conf.high), alpha=0.1, fill="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Proportion Female", y = "Predicted Mean Difference in Centile Scores")

#FIGURE 5

```{r}
# study demographics
fig5 <- ggarrange(
ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          # align = "hv", 
          common.legend = TRUE,
          legend = "bottom",
          widths=c(2, 1),
          heights=c(14, 1)),
# mean effects
ggarrange(marg.age_effects.plt,
marg.sex_effects.plt,
nrow=2, ncol=1,
align="hv",
labels=c("B", "C")),
common.legend=FALSE,
nrow=1, ncol=2,
labels=c("A", NA),
widths=c(1, 1.4)
)

@@

```{r}
# study demographics
fig5.optB <- ggarrange(
ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          # align = "hv", 
          common.legend = TRUE,
          legend = "bottom",
          widths=c(2, 1.3),
          heights=c(14, 1)),
# mean effects
ggarrange(
ggarrange(age_effects.plt, age_effects.plt.null,
          nrow=2, ncol=1, common.legend=TRUE, legend="bottom",
          align = "hv"),
ggarrange(sex_effects.plt.null, sex_effects.plt,
          nrow=2, ncol=1, common.legend=TRUE, legend="bottom",
          align = "hv"),
# align="hv",
# labels=c("B", "C")),
common.legend=FALSE,
nrow=1, ncol=2),
labels=c("A", "B"),
widths=c(1, 2.2),
nrow=1, ncol=2
)

fig5.optB_annotate <- annotate_figure(fig5.optB, top = text_grob("Figure 5. Differences in centile scores of lifespan consortium data when harmonized with ComBat-GAM and ComBatLS", color = "black", hjust = 0, x = 0))
fig5.optB_annotate
ggsave("figs/figure_5_option2.jpeg", fig5.optB_annotate, width=13, height=10, units="in")
```

### 2024-02-23 paper_main_results.Rmd
looking at CT sigmas, plots

p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
               color = "black", size = 14, hjust=.65))
p_annotate
ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
```

>>>>

p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
               color = "black", size = 14, hjust=.65))
p_annotate
ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
```
Look at relative estimates for sex in sigma across features in raw data

```{r}
gamlss.sum.df <- fread(paste0(data_path, "/ukb_basic/gamlss_summary.csv"))
gamlss.sum.df.raw <- gamlss.sum.df %>%
  dplyr::filter(dataset=="ukb-CN-data-simsites")

gamlss.sum.df.raw.sexsig <- gamlss.sum.df.raw %>%
  dplyr::filter(parameter =="sigma" & term == "sexMale") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      est_over_moment = estimate/moment.mean)

gamlss.sum.df.raw.sexsig %>%
  ggplot(aes(y=estimate, x=pheno_cat, fill=pheno_cat)) +
  geom_boxplot() +
  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
  stat_compare_means(method = "anova", label.y = 0.07)+    
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")

gamlss.sum.df.raw.sexsig %>%
  ggplot() +
  geom_histogram(aes(x=est_over_moment, fill=pheno_cat, color=pheno_cat), alpha=0.5, position="identity")

gamlss.sum.df.raw.sexsig %>%
  ggplot(aes(y=t_stat, x=pheno_cat, fill=pheno_cat)) +
  geom_boxplot() +
  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
  stat_compare_means(method = "anova", label.y = -7)+    
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")
```
